---
title: "ASDV Coursework Codes"
author: "Shola Ayeotan"
date: "2023-12-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing packages

```{r}
 # Loading necessary libraries
install.packages("caret")
install.packages("knitr")
install.packages("car")
install.packages("corrplot")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyr")
install.packages("Hmisc")
install.packages("psych")
install.packages("countrycode")
install.packages("randomForest")
install.packages("moments")
install.packages("RVAideMemoire")
install.packages("gplots")
install.packages("forecast")
install.packages("zoo")
install.packages("tseries")
install.packages("TTR")


library(gplots)
library(RVAideMemoire)
library(moments)
library(randomForest)
library(corrplot)
library(caret)
library(knitr)
library(car)
library(ggplot2)
library(dplyr)
library(tidyr)
library(Hmisc)
library(psych)
library(countrycode)
library(TTR)
library(tseries)
library(zoo)
library(forecast)
```

## **Loading the data set**

```{r}
# Reading the data set into a data frame

economic <- read.csv("economic-data.csv", sep = ",", header = TRUE)
head(economic)
```

```{r}
 str(economic)
```

## **Data Preprocessing**

```{r}
 # Renaming columns with shorter and more readable names

colnames(economic) <- c(
  "Country", "Year", "Inflation", "CPI", "GDP_Growth",
  "GDP", "Per_Capita", "Per_Capita_Growth", "Unemployment",
  "Population", "Labor_Force", "FDI_NetInflows", "FDI_Percentage",
  "Exports", "Imports", "Political_Stability", "Corruption_Control",
  "Government_Effectiveness", "Regulatory_Quality", "Rule_of_Law",    "Technology_Exports")


# Displaying the updated column names
colnames(economic)

```

```{r}
 # Extracting the names of all columns except the first two
 columns_to_convert <- names(economic)[-c(1, 2)]

 # Converting the selected columns to numeric
 economic[columns_to_convert] <- lapply(economic[columns_to_convert], as.numeric)
```

```{r}
 # Identifying missing values

 missing_values <- colSums(is.na(economic))
 missing_values
```

```{r}
# Defining a function to replace missing values

replace_missing_values <- function(
    data, column_name, country_column, missing_values) 
  {missing_row_index <- which(is.na(data[[column_name]]))
  
  countries_with_missing <- data[[country_column]][missing_row_index]
  
  for (country in unique(countries_with_missing)) {
    mean_value <- mean(data[[column_name]][data[[country_column]] == country], na.rm = TRUE)
    
  missing_in_country <- which(data[[country_column]] == country & is.na(data[[column_name]]))
    
    data[[column_name]][missing_in_country] <- mean_value}
  
  return(data)
}
```

```{r}
# Handing missing values

economic <- replace_missing_values(economic,"Inflation", "Country", missing_values)

economic <- replace_missing_values(economic, "CPI", "Country", missing_values)

economic <- replace_missing_values(economic, "Unemployment", "Country", missing_values)

head(economic)
```

```{r}
# Creating a new column called "continent"
economic$continent <- countrycode(economic$Country, origin = "country.name", destination = "continent")

# Moving the continent column to be the third in the data frame
economic <- economic[, c(1, 2, ncol(economic), 3:(ncol(economic) - 1))]

head(economic)
```

```{r}
# Visualizing continent distribution 
ggplot(economic, aes(x = continent)) +
  geom_bar() +
  labs(
    title = "Distribution of Continents",
    x = "Continent",
    y = "Count")
```

```{r}
# Checking to see number of countries in each continent
continent_counts <- economic %>%
  group_by(continent) %>%
  summarise(unique_country_count = n_distinct(Country))

print(continent_counts)
```

#### Checking for outliers

```{r}
# Creating box plots for various economic indicators

boxplot(economic$Per_Capita, main = "Per Capita", ylab = "Per Capita", col = "skyblue")

boxplot(economic$GDP, main = "Gross Domestic Product", ylab = "GDP", col = "lightgreen")

boxplot(economic$GDP_Growth, main = "Technological Exports", ylab = "Technology_Exports", col = "lightcoral")

boxplot(economic$Inflation, main = "Inflation", ylab = "Inflation", col = "gold")

boxplot(economic$Unemployment, main = "Unemployment", ylab = "Unemployment", col = "lightblue")

boxplot(economic$CPI, main = "Consumer Price Index", ylab = "CPI", col = "plum")

boxplot(economic$Population, main = "Political_Stability", ylab = "Population", col = "lightpink")

boxplot(economic$Labor_Force, main = "Labor Force", ylab = "Labor Force", col = "lightsteelblue")
```

## **Statistical Analysis**

```{r}
describe(economic)
```

```{r}
# Creating a function to summarize and visualize statistics for each continent
visualizeStats <- function(data, variable_name) {
  descriptive_stats <- data %>%
    dplyr::group_by(continent) %>%
    dplyr::summarize(
      mean_value = mean(!!sym(variable_name), na.rm = TRUE),
      median_value = median(!!sym(variable_name), na.rm = TRUE),
      sd_value = sd(!!sym(variable_name), na.rm = TRUE),
      skewness_value = skewness(!!sym(variable_name), na.rm = TRUE))

  # Display the results
  print(descriptive_stats)

  # Visualizing the statistics
  bar_plot <- ggplot(descriptive_stats, aes(x = continent)) +
    geom_bar(aes(y = mean_value), fill = "skyblue", stat = "identity") +
    labs(title = paste("Mean", variable_name, "across continents"), x = "Continent", y = "Mean Value")

  sd_plot <- ggplot(descriptive_stats, aes(x = continent)) +
    geom_bar(aes(y = sd_value), fill = "coral", stat = "identity") +
    labs(title = paste("Standard Deviation of", variable_name, "Across Continents"), x = "Continent", y = "SD Value")

  density_plot <- ggplot(data, aes(x = .data[[variable_name]])) +
    geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black") +
    geom_density(alpha = 0.5, fill = "orange") +
    geom_vline(aes(xintercept = mean(.data[[variable_name]])), color = "blue", linetype = "dashed", size = 1) +
    geom_vline(aes(xintercept = median(.data[[variable_name]])), color = "red", linetype = "dashed", size = 1) +
    labs(title = paste("Distribution of", variable_name, "with Skewness"),
         x = paste(variable_name, "(current US$)"),
         y = "Density")

  box_plot <- ggplot(data, aes(x = continent)) +
    geom_boxplot(aes(y = .data[[variable_name]])) +
    labs(title = "Distribution Across Continents", x = "Continent", y = "Value")

  print(bar_plot)
  print(sd_plot)
  print(density_plot)
  print(box_plot)
}
```

```{r}
# Initializing the function on GDP
visualizeStats(economic, "GDP")
```

```{r}
# Initializing the function on GDP_Growth

visualizeStats(economic, "GDP_Growth")
```

```{r}
visualizeStats(economic, "Inflation")
```

```{r}
visualizeStats(economic, "Unemployment")
```

```{r}
visualizeStats(economic, "Population")
```

```{r}
visualizeStats(economic, "Political_Stability")
```

```{r}
visualizeStats(economic, "Corruption_Control")
```

```{r}
visualizeStats(economic, "Government_Effectiveness")
```

### **Exploratory Data Analysis (EDA)**

```{r}
# Visualizing GDP across the years

ggplot(economic, aes(x = Year, y = GDP_Growth, fill = continent)) +
  geom_area() +
  labs(title = "Gross Domestic Product", x = "Year", y = "GDP (current US$)")
```

```{r}
# Visualizing Inflation levels

ggplot(economic, aes(x = Year, y = Inflation, fill = continent)) +
  geom_area() +
  labs(title = "Inflation Across Continents Over Years", x = "Year", y = "Inflation")
```

```{r}
# Visualizing Consumer Price Index

ggplot(economic, aes(x = Year, y = CPI, fill = continent)) +
  geom_area() +
  labs(title = "CPI Across Continents Over Years", x = "Year", y = "Inflation")
```

```{r}
# Visualizing continental unemployment levels

ggplot(economic, aes(x = Year, y = Unemployment, fill = continent)) +
  geom_area() +
  labs(title = "Unemployment", x = "Year", y = "Unemployment")
```

```{r}
# Visualizing country unemployment levels

ggplot(economic, aes(x=Country, y=Unemployment)) +
  geom_point(size=3) +
  geom_segment(aes(x=Country,
                   xend=Country,
                   y=0,
                   yend=Unemployment)) +
  labs(title="Unemployment Figures",
       subtitle="Youth Unemployment Rise for each country",
       caption="Youth Unemployment Rise") +
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

```{r}
# Visualizing Population numbers

ggplot(economic, aes(x = Year, y = Population, fill = continent)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Population",
    x = "Year",
    y = "Population")
```

```{r}
# High technology exports across continents

ggplot(economic, aes(x = Year, y = Technology_Exports, fill = continent)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Technology Exports Across Continents Over Years",
    x = "Year",
    y = "Technology Export")

ggplot(economic, aes(x = Technology_Exports, y = Country, color = Year)) +
  geom_point() +
  labs(
    title = "Technology Exports Across Countries Over Years",
    x = "Technology Export (in Billions USD)",
    y = "Year"
  )
```

```{r}
# Foreign Direct Investment 

ggplot(economic, aes(x = Year, y = FDI_NetInflows, fill = continent)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Foreign Direct Investment (% of GDP)", x = "Year", y = "FDI")
```

```{r}
# Political Stability and Absence of Violence/Terrorism

ggplot(economic, aes(x = Year, y = `Political_Stability`, group = `continent`, color = `continent`)) +
  geom_boxplot() +
  labs(title = "Political Stability over time",
       x = "Year",
       y = "Political Stability")
```

### **Checking for Normal Distribution**

```{r}
# Function to create histograms and Q-Q plots

check_normality <- function(variable_name) {
  p1 <- ggplot(economic, aes(x = .data[[variable_name]])) +
    geom_histogram(binwidth = 10, fill = "blue", color = "black", alpha = 0.7) +
    labs(title = paste("Histogram for", variable_name),
         x = variable_name,
         y = "Frequency") +
    theme_minimal()

  p2 <- ggplot(economic, aes(x = .data[[variable_name]])) +
    geom_density(fill = "blue", alpha = 0.7) +
    labs(title = paste("Density Plot for", variable_name),
         x = variable_name,
         y = "Density") +
    theme_minimal()

  print(p1)
  print(p2)
}
```

```{r}
check_normality("CPI")
```

```{r}
check_normality("Inflation")
```

```{r}
check_normality("GDP_Growth")
```

```{r}
check_normality("Unemployment")
```

### **Shapiro--Wilk test**

```{r}
# Function to perform Shapiro-Wilk test

shapiro_test_all <- function(data) {
  cat("Shapiro-Wilk Test Results\n\n")

  for (variable_name in colnames(data)) {
    if (is.numeric(data[[variable_name]])) {
      shapiro_test_result <- shapiro.test(data[[variable_name]])
      cat("Variable:", variable_name, "\n")
      cat("  W =", shapiro_test_result$statistic, ", p-value =", shapiro_test_result$p.value, "\n")

      # Check if we should reject or accept the null hypothesis
      if (shapiro_test_result$p.value > 0.05) {
        cat("  Result: Do not reject the null hypothesis (data may be normally distributed)\n\n")
      } else {
        cat("  Result: Reject the null hypothesis (data may not be normally distributed)\n\n")
      }
    }
  }
}
```

```{r}
shapiro_test_all(economic)
```

**Observation:**

The P-value of all the variables are less than the significance level 0.05, hence we proceed to **reject the Null Hypothesis** and conclude that the data does not follow a normal distribution.

### **Conducting Logarithmic** **Transformations**

```{r}
economic_log <- economic

# Selecting numeric columns for transformation
numeric_columns <- sapply(economic_log, is.numeric)

# Applying logarithmic transformation to the numeric columns
economic_log[numeric_columns] <- lapply(economic_log[numeric_columns], function(x) log(x + 1))

# Checking the distributions after transformation
par(mfrow = c(2, 2))
for (col in names(economic_log[numeric_columns])) {
  hist(economic_log[[col]], main = col, col = "lightblue")
}
```

```{r}
 # Identifying missing values

 missing_values <- colSums(is.na(economic_log))
 missing_values
```

```{r}
## Handing missing values

economic_log <- replace_missing_values(economic_log, "Inflation", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "GDP_Growth", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Per_Capita_Growth", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "FDI_NetInflows", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "FDI_Percentage", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Political_Stability", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Corruption_Control", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Government_Effectiveness", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Regulatory_Quality", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Rule_of_Law", "Country", missing_values)

economic_log <- replace_missing_values(economic_log, "Rule_of_Law", "Country", missing_values)
```

```{r}
missing_values <- economic_log[is.na(economic_log$FDI_NetInflows) | is.na(economic_log$Political_Stability), c("Country", "FDI_NetInflows", "Political_Stability")]

# Print the result
print(missing_values)
```

## **Correlation Analysis**

```{r}
# Creating the Spearman correlation matrix
spearman_cor  <- cor(economic[, 4:ncol(economic)], method="spearman")

corrplot(spearman_cor , method = "color")
```

#### Correlation within different indicator groupings

```{r}
 # Selecting governance-related variables
 govt_var <- economic[, c("GDP", "GDP_Growth", "Per_Capita", "Per_Capita_Growth", "Political_Stability",
 "Corruption_Control", "Government_Effectiveness", "Regulatory_Quality", "Rule_of_Law")]

 # Computing the correlation matrix
govt_matrix <- cor(govt_var, method="spearman")

corrplot(cor(govt_matrix), method = "number", type = "upper")
```

```{r}
# Selecting trade-related variables
trade_var <- economic[, c("GDP", "GDP_Growth", "Per_Capita", "Per_Capita_Growth", "Imports","Exports", "Technology_Exports")]

# Computing the correlation matrix
trade_matrix <- cor(trade_var, method="spearman")

corrplot(cor(trade_var), method = "number", type = "upper")
```

```{r}
# Selecting inflation-related variables
inflation_var <- economic[, c("GDP", "GDP_Growth", "Per_Capita", "Per_Capita_Growth", "Inflation", "CPI")]

# Computing the correlation matrix
inflation_matrix <- cor(inflation_var, method="spearman")

corrplot(cor(inflation_var), method = "number", type = "upper")
```

```{r}
# Selecting labour-related variables
labour_var <- economic[, c("GDP", "GDP_Growth", "Per_Capita", "Per_Capita_Growth", "Unemployment", "Population", "Labor_Force")]

# Computing the correlation matrix
labour_matrix <- cor(labour_var, method="spearman")

corrplot(cor(labour_var), method = "number", type = "upper")
```

```{r}
# Selecting foreign investment variables
investment_var <- economic[, c("GDP", "GDP_Growth", "Per_Capita", "Per_Capita_Growth", "FDI_NetInflows", "FDI_Percentage")]

# Compute correlation matrix
investment_cor_matrix <- cor(investment_var, method="spearman")

corrplot(cor(investment_var), method = "number", type = "upper")
```

## **Hypothesis Testing**

### Testing with Correlation

**Hypothesis 1**

```{r}
# Testing the correlation between Inflation and Unemployment 

cor_result <- cor.test(economic$Inflation, economic$Unemployment, method = "spearman", exact = FALSE)
print(cor_result)

# Decision Rule
if (cor_result$p.value < 0.05) 
  {cat("Reject the null hypothesis. There is no evidence of a negative correlation.")} else {
  cat("Fail to reject the null hypothesis. There is enough evidence of a negative correlation.")}
```

**Hypothesis 2**

```{r}
cor_result <- cor.test(economic$FDI_NetInflows, economic$GDP, method = "spearman", exact = FALSE)

print(cor_result)

# Decision Rule
if (cor_result$p.value < 0.05) 
  {cat("Reject the null hypothesis. There is significant correlation.")} else {
  cat("Fail to reject the null hypothesis. There is no evidence of a correlation.")}
```

**Hypothesis 3**

```{r}
cor_result <- cor.test(economic$Government_Effectiveness, economic$Regulatory_Quality, 
                       method = "spearman", exact = FALSE)

print(cor_result)

# Decision Rule
if (cor_result$p.value < 0.05) 
  {cat("Reject the null hypothesis. There is significant correlation.")} else {
  cat("Fail to reject the null hypothesis. There is no evidence of a correlation.")}
```

**Hypothesis 4**

```{r}
cor_result <- cor.test(economic$GDP, economic$Exports, method = "spearman", exact = FALSE)

print(cor_result)

# Decision Rule
if (cor_result$p.value < 0.05) 
  {cat("Reject the null hypothesis. There is significant correlation.")} else {
  cat("Fail to reject the null hypothesis. There is no evidence of a correlation.")}
```

### **Testing using ANOVA**

```{r}
# Randomly selecting a country from each continent

set.seed(300)

selected_countries <- economic_log %>%
  group_by(continent) %>%
  slice_sample(n = 2) %>%
  ungroup()

hypothesis <- economic[economic$Country %in% 
                         selected_countries$Country, ]

print(selected_countries)
```

#### **Hypothesis 5: Mean** Inflation

```{r}
# Checking for outlines
boxplot(Inflation ~ continent, data=hypothesis, xlab="Continents", 
        ylab="Inflation", main="Inflation Across continents")
```

```{r}
ggplot(hypothesis, aes(x = continent, y = Inflation, fill = continent)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Mean Inflation Across Continents",
       x = "Continent",
       y = "Mean inflation") +
  theme_minimal()
```

```{r}
# Checking for normal distribution
byf.shapiro(Inflation  ~ continent, data=hypothesis)
```

```{r}
 # Checking for homogeneity of variances

bartlett.test(Inflation ~ Country, data=hypothesis)
```

```{r}
# Conducting a post-hoc with Welch

oneway.test(Inflation ~ Country,data=hypothesis, var.equal = FALSE)
```

```{r}
# Running a Post-Hoc Test

inflation_anova <- aov(Inflation ~ continent, data = hypothesis)
TukeyHSD(inflation_anova, which = "continent")
```

#### **Hypothesis 6: Mean** GDP

```{r}
# Checking for outliers

boxplot(GDP ~ continent, data=hypothesis,xlab="Continents", 
        ylab="GDP",main="GDP across continents")
```

```{r}
ggplot(hypothesis, aes(x = continent, y = GDP, fill = continent)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Mean GDP Across Continents",
       x = "Continent",
       y = "Mean GDP") +
  theme_minimal()
```

```{r}
# Checking for normal distribution
byf.shapiro(GDP ~ continent, data=hypothesis)
```

```{r}
 # Checking for homogeneity of variances
bartlett.test(GDP ~ Country, data=hypothesis)
```

```{r}
# Conducting a post-hoc with Welch
oneway.test(GDP ~ Country,data=hypothesis, var.equal = FALSE)
```

```{r}
# Running a Post-Hoc Test

GDP_anova <- aov(GDP ~ continent, data = hypothesis)
TukeyHSD(GDP_anova, which = "continent")
```

#### **Hypothesis 7: Mean** Unemployment

```{r}
# Checking for outliers

boxplot(Unemployment ~ continent, data=hypothesis, xlab="Continents", 
        ylab="Unemployment", main="Unemployment across continents")
```

```{r}
ggplot(hypothesis, aes(x = continent, y = Unemployment, fill = continent)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Mean Unemployment Across Continents",
       x = "Continent",
       y = "Mean Unemployment") +
  theme_minimal()
```

```{r}
# Checking for normal distribution
byf.shapiro(Unemployment ~ continent, data=hypothesis)
```

```{r}
 # Checking for homogeneity of variances
bartlett.test(Unemployment ~ Country, data=hypothesis)
```

```{r}
# Conducting a posthoc with Welch
oneway.test(Unemployment ~ Country,data=hypothesis, var.equal = FALSE)
```

```{r}
# Running a Post-Hoc Test

Unemployment_anova <- aov(Unemployment ~ continent, data = hypothesis)
TukeyHSD(Unemployment_anova, which = "continent")
```

## **Regression Analysis**

### Simple Linear Regression: Trade Balances vs GDP

```{r}
# Randomly selecting a country from each continent

set.seed(200)

selected_countries <- economic_log %>%
  group_by(continent) %>%
  slice_sample(n = 1) %>%
  ungroup()

print(selected_countries)
```

```{r}
# Creating a data subset for the selected countries using our transformed data

economicIndicators <- economic_log[
  economic_log$Country %in% selected_countries$Country,
  c("Country", "continent", "Per_Capita", "GDP", "Labor_Force", "Exports", 
    "Imports", "Technology_Exports")]

economicIndicators_Africa <- economicIndicators[economicIndicators$continent == "Africa", ]
economicIndicators_Americas <- economicIndicators[economicIndicators$continent == "Americas", ]
economicIndicators_Asia <- economicIndicators[economicIndicators$continent == "Asia", ]
economicIndicators_Europe <- economicIndicators[economicIndicators$continent == "Europe", ]
```

```{r}

```

#### **Continent-Specific SLR Models**

**Africa**

```{r}
# Fitting an SLR for a single country in Africa
ecoModel1 <- lm(GDP ~ Exports, data = economicIndicators_Africa)
print(summary(ecoModel1))


# Plotting the regression line
plot( GDP ~ Exports, data = economicIndicators_Africa,
 col = "blue",
 main = "Regression: GDP & Exports for Africa",
 xlab = "Exports",
 ylab = "GDP ($billion)")

abline(ecoModel1, col="red")



## Checking SLR assumptions
plot(ecoModel1, 1)

plot(ecoModel1, 2)

plot(ecoModel1, 3)
```

**Americas**

```{r}
# Fitting an SLR for a single country in Americas
ecoModel2 <- lm(GDP ~ Exports, data = economicIndicators_Americas)
print(summary(ecoModel2))


# Plotting the regression line
plot( GDP ~ Exports, data = economicIndicators_Americas,
col = "blue",
 main = "Regression: GDP & Exports for Americas",
 xlab = "Exports",
 ylab = "GDP ($billion)")

abline(ecoModel2, col="red")


## Checking SLR assumptions
plot(ecoModel2, 1)
plot(ecoModel2, 2)
plot(ecoModel2, 3)
```

**Asia**

```{r}
# Fitting an SLR for a single country in Asia
ecoModel3 <- lm(GDP ~ Exports, data = economicIndicators_Asia)
print(summary(ecoModel3))

# Plotting the regression line
plot( GDP ~ Exports, data = economicIndicators_Asia,
col = "blue",
 main = "Regression: GDP & Exports for Asia",
 xlab = "Exports",
 ylab = "GDP ($billion)")

abline(ecoModel3, col="red")


## Checking SLR assumptions
plot(ecoModel3, 1)
plot(ecoModel3, 2)
plot(ecoModel3, 3)
```

**Europe**

```{r}
ecoModel4 <- lm(GDP ~ Exports, data = economicIndicators_Europe)
print(summary(ecoModel4))


# Plotting the regression line
plot( GDP ~ Exports, data = economicIndicators_Europe,
col = "blue",
 main = "Regression: GDP & Exports for Europe",
 xlab = "Exports",
 ylab = "GDP ($billion)")

abline(ecoModel4, col="red")

## Checking SLR assumptions
plot(ecoModel4, 1)
plot(ecoModel4, 2)
plot(ecoModel4, 3)
```

#### Generalized SLR Model

(Based on all countries)

```{r}
ecoModel5 <- lm(GDP ~ Exports, data = economic_log)
print(summary(ecoModel5))


# Plotting the regression line
plot( GDP ~ Exports, data = economic_log,
 col = "blue",
 main = "Regression: GDP & Exports",
 xlab = "Exports",
 ylab = "GDP ($billion)")

abline(ecoModel5, col="red")
```

```{r}
## Checking all regression assumptions
plot(ecoModel5, 1)

plot(ecoModel5, 2)

residuals <- residuals(ecoModel5)
hist(residuals, main = "Histogram of Residuals", col = "red")

plot(ecoModel5, 3)
```

**Predicting with the SLR model**

```{r}
# Creating new data for prediction
pred_data = c(290551190, 303828107, 535316688)  

# Setting coefficients
intercept = 3.81283
exports_coefficient = 0.90416

# Making predictions
predicted_GDP = intercept + exports_coefficient * new_exports

print(predicted_GDP)
```

#### Generalized MLR Model

```{r}
# Fitting a Multiple Linear Regression model
ecoModel6 <- lm(GDP ~ Exports + Imports , data = economic_log)
print(summary(ecoModel6))

# Adding an extra independent variable to the model
ecoModel7 <- lm(GDP ~ Exports + Imports + Technology_Exports   , data = economic_log)
print(summary(ecoModel7))

# Adding Imports to the model
ecoModel8 <- lm(GDP ~ Exports + Imports + Labor_Force, data = economic_log)
print(summary(ecoModel8))
```

```{r}
# Plotting the regression line
ggplot(ecoModel7, aes(x = Exports, y = GDP)) +
  geom_point(aes(color = Imports)) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "blue") +
  labs(title = "Multiple Linear Regression",
       x = "Exports",
       y = "Gross Domestic Product",
       color = "Imports") +
  theme_minimal()
```

```{r}
## Checking SLR assumptions

plot(ecoModel7, 1)
plot(ecoModel7, 2)
plot(ecoModel7, 3)
```

**Conducting Predictions based on the MLR model**

```{r}

intercept = -2.46763
exports_coefficient = -0.24459
imports_coefficient = 1.46749
tech_exports_coefficient = -0.09584


# Creating Sample data
exports_value = 807060793372  
imports_value = 862311688153   
tech_exports = 67496532150 

# Making predictions
predicted_GDP = intercept + 
                exports_coefficient * exports_value + 
                imports_coefficient * imports_value + 
                tech_exports_coefficient * tech_exports

print(predicted_GDP)

```

### **Multiple Linear Regression:** Government Indices vs GDP Per Capita

```{r}
# Selecting governance-related variables
social_var <- economic_log[, c("Per_Capita", "Corruption_Control", "Political_Stability", "Government_Effectiveness", "Regulatory_Quality", "Rule_of_Law")]

# Computing correlation matrix
social_matrix <- cor(social_var, method="spearman")
corrplot(cor(social_var), method = "number", type = "upper")
```

```{r}
pairs(social_var, lower.panel = NULL, pch = 19,cex = 0.2)
```

```{r}
# Fitting the Multiple Linear Regression model
socModel1 <- lm(Per_Capita ~ Government_Effectiveness + 
                  Corruption_Control, data = economic_log)
print(summary(socModel1))



# Adding an extra independent variable to the mix
socModel2 <- lm(Per_Capita ~ Government_Effectiveness + 
                  Corruption_Control + Political_Stability, data = economic_log)
print(summary(socModel2))



# Adding Regulatory Quality to the mix
socModel3 <- lm(Per_Capita ~ Government_Effectiveness + Corruption_Control 
                + Political_Stability + Regulatory_Quality, data = economic_log)
print(summary(socModel3))
```

```{r}
# Plotting the regression line

ggplot(socModel3, aes(x = Government_Effectiveness, y = Per_Capita)) +
  geom_point(aes(color = Regulatory_Quality)) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "blue") +
  labs(title = "Multiple Linear Regression",
       x = "Government Effectiveness",
       y = "GDP Per Capita",
       color = "Regulatory Quality") +
  theme_minimal()
```

```{r}
## Checking all regression assumptions
plot(socModel3, 1)

plot(socModel3, 2)

residuals <- residuals(socModel3)
hist(residuals, main = "Histogram of Residuals", col = "lightblue")

plot(socModel3, 3)
```

**Conducting Predictions based on the MLR model**

```{r}

# Predicting Per Capita income based on the model

intercept = 9.14689
govteffectiveness_coeff = 0.79609
corruption_control_coeff = -0.14349
political_stability_coeff = 0.24382
regulatory_quality_coeff = 0.67249


# Creating Sample data for 3 years
pred_data <- data.frame(
  Government_Effectiveness = c(-1.252720952, 1.398337722, 0.996483803),
  Corruption_Control = c(1.038217425, 1.437633395, 1.260388613),
  Political_Stability = c(-0.01487787627, 0.3860425651, 0.1178766638),
  Regulatory_Quality = c(0.930574059, 1.238437533, 0.996483803))


# Making predictions
pred_PerCapita <- predict(socModel3, newdata = pred_data)

print(pred_PerCapita)

```

## **Time Series Analysis**

```{r}
# Randomly selecting countries from each continent

set.seed(20)

selected_countries <- economic_log %>%
  group_by(continent) %>%
  slice_sample(n = 1) %>% 
  ungroup()

print(selected_countries)
```

```{r}
# Creating a data subset for the selected countries

GDP_Data <- economic_log[economic_log$Country %in% selected_countries$Country, 
                         c("Country", "continent", "GDP", "GDP_Growth")]

GDP_Data_Africa <- GDP_Data[GDP_Data$continent == "Africa", ]
GDP_Data_Americas <- GDP_Data[GDP_Data$continent == "Americas", ]
GDP_Data_Asia <- GDP_Data[GDP_Data$continent == "Asia", ]
GDP_Data_Europe <- GDP_Data[GDP_Data$continent == "Europe", ]
```

### Forecasting with Holt-Winters

##### **AFRICA (ETHIOPIA) FORECAST**

```{r}
# Creating a time series object for yearly data
seriesAfrica <- ts(GDP_Data_Africa$GDP_Growth, start = c(2008), end = c(2022))
plot.ts(seriesAfrica)
```

```{r}
# Calculate and comparing Simple Moving Average (SMA) with different sizes 
GDPseriesSMA3 <- SMA(seriesAfrica, n=3)
plot.ts(GDPseriesSMA3)

GDPseriesSMA8 <- SMA(seriesAfrica, n=8)
plot.ts(GDPseriesSMA8)
```

```{r}
# Using Holt-Winters Exponential Smoothing (with trend component)

GDPseries_smoothing2 <-  HoltWinters(seriesAfrica, gamma=FALSE)
GDPseries_smoothing2
plot(GDPseries_smoothing2)
```

```{r}
# Computing the Sum of Squared Errors (SSE)
GDPseries_smoothing2$SSE
```

```{r}
# Forecasting the next 15  years 
GDPseries_forecast2 <- forecast(GDPseries_smoothing2, h=15)
GDPseries_forecast2
plot(GDPseries_forecast2)
```

```{r}
# Plotting autocorrelation function
acf(GDPseries_forecast2$residuals, lag.max=20 , na.action = na.pass)

# Performing the Ljung-Box test 
Box.test(GDPseries_forecast2$residuals, lag=10, type="Ljung-Box")
```

```{r}
# Defining function for forecast errors

plotForecastErrors <- function(forecasterrors)
{
  # make a histogram of the forecast errors:
  mybinsize <- IQR(forecasterrors,na.rm=TRUE)/4
  mysd   <- sd(forecasterrors,na.rm=TRUE)
  mymin  <- min(forecasterrors,na.rm=TRUE) - mysd*5
  mymax  <- max(forecasterrors,na.rm=TRUE) + mysd*3
  # generate normally distributed data with mean 0 and standard deviation mysd
  mynorm <- rnorm(10000, mean=0, sd=mysd)
  mymin2 <- min(mynorm)
  mymax2 <- max(mynorm)
  if (mymin2 < mymin) { mymin <- mymin2 }
  if (mymax2 > mymax) { mymax <- mymax2 }
  # make a red histogram of the forecast errors, with the normally distributed data overlaid:
  mybins <- seq(mymin, mymax, mybinsize)
  hist(forecasterrors, col="red", freq=FALSE, breaks=mybins)
  # freq=FALSE ensures the area under the histogram = 1
  # generate normally distributed data with mean 0 and standard deviation mysd
  myhist <- hist(mynorm, plot=FALSE, breaks=mybins)
  # plot the normal curve as a blue line on top of the histogram of forecast errors:
  points(myhist$mids, myhist$density, type="l", col="blue", lwd=2)
}
```

```{r}
# Plotting the time series of residuals and forecast errors

plot.ts(GDPseries_forecast2$residuals)

GDPseries_forecast2$residuals <-GDPseries_forecast2$residuals[!is.na(GDPseries_forecast2$residuals)]
plotForecastErrors(GDPseries_forecast2$residuals)
```

##### **AMERICAS (MEXICO) FORECAST**

```{r}
seriesAmericas <- ts(GDP_Data_Americas$GDP_Growth, start = c(2008), end = c(2022))
plot.ts(seriesAmericas)
```

```{r}
GDPseries_smoothing <-  HoltWinters(seriesAmericas, gamma=FALSE)
GDPseries_smoothing
```

```{r}
GDPseries_smoothing$SSE
plot(GDPseries_smoothing)
```

```{r}
GDPseries_forecast <- forecast(GDPseries_smoothing, h=19)
GDPseries_forecast
plot(GDPseries_forecast)
```

```{r}

acf(GDPseries_forecast$residuals, lag.max=20 , na.action = na.pass)

Box.test(GDPseries_forecast$residuals, lag=20, type="Ljung-Box")
```

```{r}
plot.ts(GDPseries_forecast$residuals)


GDPseries_forecast$residuals <-GDPseries_forecast$residuals[!is.na(GDPseries_forecast$residuals)]
plotForecastErrors(GDPseries_forecast$residuals)
```

##### **ASIA (CHINA) FORECAST**

```{r}
seriesAsia <- ts(GDP_Data_Asia$GDP_Growth, start = c(2008), end = c(2022))
plot.ts(seriesAsia)
```

```{r}
GDPseries_smoothing3 <-  HoltWinters(seriesAsia, gamma=FALSE)
GDPseries_smoothing3
```

```{r}
GDPseries_smoothing3$SSE
plot(GDPseries_smoothing3)
```

```{r}
GDPseries_forecast3 <- forecast(GDPseries_smoothing3, h=5)
GDPseries_forecast3
plot(GDPseries_forecast3)
```

```{r}

acf(GDPseries_forecast3$residuals, lag.max=20 , na.action = na.pass)
Box.test(GDPseries_forecast3$residuals, lag=20, type="Ljung-Box")
```

```{r}
plot.ts(GDPseries_forecast3$residuals)


GDPseries_forecast3$residuals <-GDPseries_forecast3$residuals[!is.na(GDPseries_forecast3$residuals)]
plotForecastErrors(GDPseries_forecast3$residuals)
```

##### **EUROPE (GERMANY) FORECAST**

```{r}
seriesEurope <- ts(GDP_Data_Europe$GDP_Growth, start = c(2008), end = c(2022))
plot.ts(seriesEurope)
```

```{r}
GDPseries_smoothing4 <-  HoltWinters(seriesEurope, gamma=FALSE)
GDPseries_smoothing4
```

```{r}
GDPseries_smoothing4$SSE
plot(GDPseries_smoothing4)
```

```{r}
GDPseries_forecast4 <- forecast(GDPseries_smoothing4, h=19)
GDPseries_forecast4
plot(GDPseries_forecast4)
```

```{r}
acf(GDPseries_forecast4$residuals, lag.max=20 , na.action = na.pass)
Box.test(GDPseries_forecast4$residuals, lag=20, type="Ljung-Box")
```

```{r}
plot.ts(GDPseries_forecast4$residuals)

GDPseries_forecast4$residuals <-GDPseries_forecast4$residuals[!is.na(GDPseries_forecast4$residuals)]
plotForecastErrors(GDPseries_forecast4$residuals)
```

### Forecasting with ARIMA Model

```{r}
# Differencing the time series
seriesAfrica_diff1 <- diff(seriesAfrica, differences=1)
plot.ts(seriesAfrica_diff1)
```

```{r}
# Exploring stationarity with two differences
seriesAfrica_diff2 <- diff(seriesAfrica, differences=2)
plot.ts(seriesAfrica_diff2)
```

```{r}
# Finding values of p and q

# Plotting ACF
acf(seriesAfrica_diff2, lag.max=20)
acf(seriesAfrica_diff2, lag.max=20, plot=FALSE)
```

```{r}
# Plotting PACF
pacf(seriesAfrica_diff2, lag.max=20)
pacf(seriesAfrica_diff2, lag.max=20, plot=FALSE)
```

```{r}
# Fitting the ARIMA model
seriesAfrica_arima <- arima(seriesAfrica, order=c(0,1,1))
seriesAfrica_arima
```

```{r}
# Generating forecasts using the ARIMA model
seriesAfrica_forecasts <- forecast(seriesAfrica_arima, h=5)
seriesAfrica_forecasts
plot(seriesAfrica_forecasts)
```

```{r}
acf(seriesAfrica_forecasts$residuals, lag.max=20)
Box.test(seriesAfrica_forecasts$residuals, lag=14, type="Ljung-Box")
```

```{r}
plot.ts(seriesAfrica_forecasts$residuals)

plotForecastErrors(seriesAfrica_forecasts$residuals)
```
